%==============================================================================
% ABSTRACT (NBR 6028 — versão em inglês do resumo)
% O ambiente resumo[Abstract] já gerencia o otherlanguage internamente no abntex2
%==============================================================================

\begin{resumo}[Abstract]

This book offers a complete and self-taught approach to the intersection between Dentistry and Artificial Intelligence. Organized in 5 parts and 28 chapters, totaling approximately 500 pages, the work guides the reader from level zero (layperson) to PhD/developer level. Part I presents the historical foundations of AI in dentistry, introduction to machine learning, and Python with Google Colab. Part II deepens the theory of convolutional neural networks, semantic segmentation, dental radiology with deep learning, and classification of oral lesions and oral cancer. Part III establishes the methodology of dental software engineering with Spec-Driven Development (SDD) and Test-Driven Development (TDD), including guided projects, model interpretability, code auditing, and scientific reproducibility. Part IV explores advanced applications in periodontics, implantology, orthodontics, digital twins, foundation models, and dental RAG. Part V addresses scientific publication, ethics, LGPD, regulation, dental education with AI, and future perspectives. Each chapter combines rigorous bibliographic review with auditable citations (active DOI), practice guided by specification (SDD) and tests (TDD) in Python and Google Colab. The work includes over 100 referenced scientific articles, 93 automated tests, critical reviews, and public dental datasets.

\textbf{Keywords:} Dentistry. Artificial Intelligence. Deep Learning. Image Processing. Digital Twins. SDD. TDD. Python. Google Colab.
\end{resumo}
